Two applications of logic programming to Coq

### 2021-03-05 papers/types2020/revisions/intro.tex
Updated intro with mild changes.

Recently, Enrico Tassi et al.\ developed the Elpi
implementation~\cite{dunchev15lpar} of \lP~\cite{miller12proghol} and
more recently, Tassi has made Elpi available as the Coq-Elpi
plugin~\cite{tassi18coqpl} (\url{https://github.com/LPCIC/coq-elpi})
to the Coq proof assistant.  This implementation of \lP extends
earlier ones in primarily two directions: First, Elpi adds a notion of
constraints and constraints handling rules, which makes it more
expressive than the Teyjus implementation~\cite{nadathur99cade} of
\lP. Second, the plugin version of Elpi comes equipped with a
quotation and anti-quotation syntax for mixing Coq expressions with
\lP program elements and an API for accessing the Coq environment,
type checking included.

@@

In this paper, we present two applications of Elpi within Coq. With
these examples, we shall illustrate that Elpi is useful not only
because of its meta-programming features, but also because it soundly
implements a higher-order intuitionistic logic: such implementations
of higher-order logic have long been known to provide powerful and
flexible approaches to implementing many different logics and their
proof systems~\cite{felty89phd,felty88cade,paulson89jar}.  Following
that tradition, the Elpi system makes it possible to encode the proofs
and proof theory of various subsets of the logic behind Coq (see also~\cite{felty93lics}).

>>>>

Recently, Enrico Tassi et al.\ developed the Elpi
implementation~\cite{dunchev15lpar} of \lP~\cite{miller12proghol} and
more recently, Tassi has made Elpi available as the Coq-Elpi
plugin~\cite{tassi18coqpl} (\url{https://github.com/LPCIC/coq-elpi})
to the Coq proof assistant.  This implementation of \lP extends
earlier ones in primarily two directions: First, Elpi adds a notion of
constraints and constraints handling rules, which makes it more
expressive than the Teyjus implementation~\cite{nadathur99cade} of
\lP. Second, the plugin version of Elpi comes equipped with a
quotation and anti-quotation syntax for mixing Coq expressions with
\lP program elements and an API for accessing the Coq environment,
including its type checker.

@@

In this paper, we present two applications of Elpi within Coq. With
these applications, we shall illustrate that Elpi is useful not only
because of its meta-programming features but also because it soundly
implements a higher-order intuitionistic logic: such implementations
of higher-order logic have long been known to provide powerful and
flexible approaches to implementing many different logics and their
proof systems~\cite{felty89phd,felty88cade,paulson89jar}.  Following
that tradition, the Elpi system makes it possible to encode the proofs
and proof theory of various subsets of the logic behind Coq 
(see also~\cite{felty93lics}).
